\TITLE

### 2020-12-19 response-v3-content.tex

\definecolor{PROGRESS}{cmyk}{0.10, 0.10, 0.0, 0.0, 1.00}
\definecolor{TODO}{cmyk}{0.0, 0.10, 0.10, 0.00, 1.00}
\definecolor{DONE}{cmyk}{0.10, 0.0, 0.10, 0.0, 1.0}

\newcommand{\REVIEW}[3]{
  \begin{tcolorbox}[colback={#1}]
    {\bf {#1}. {#2}}
 \end{tcolorbox}
  \begin{quote}
    {#3}
  \end{quote}
}

@@

REVISION DUE: 01.12.2021

\section{Additional actions taken}

\REVIEW{DONE}{Using the JDS template} {We used the new JDS \LaTeX
  template available from the Journal page instead of the older one
  that we used. This also includes The change of citations in brackets
  when cited. URLs have been added with the standard urlbst tool so
  they can be used in the refernces. The jdsurl.bst is included in the
  source.}

@@

\section*{Reviewer 1}

My comments are the following:

The revised manuscript is much smoother than the first version, but I
still feel that less than sufficient care was put into the
writing. The reviewer pointed some in the pdf report. I have some
additional editorial suggestions:

\bigskip

  \REVIEW{PROGRESS} {Submit the point-by-point response to the
    reviewer's comments and my comments in a pdf file.}  {The original
    submisison web page had a field to enter the comments as text
    which we used. We will change this and introduce a PDF file
    instead, which is easier for us also.}

  \REVIEW{PROGRESS} {Submit the point-by-point response to the
    reviewer's comments and my comments in a pdf file.}
  {The original submisison web page had a field to enter the comments
    as text which we used. We will change this and introduce a PDF
    file instead, which is easier for us also.}

  \REVIEW{DONE} {Needs page number (line number) for ease of
    referencing in the review.}
  {Page numbers and line numbers were
    activated with the review style from the new JDS \LaTeX{}
    template}

  \REVIEW{DONE} {Do not italicize in the text.}
  {All italicized text
    have been removed.  Formulas enclosed in dollar signs are in
    italized text as is default for the JDS document style.}

  \REVIEW{DONE} {AICov needs to be defined in the abstract.}
  {We added ``We present the architecture of an artificial
    intelligence enhanced COVID-19 analysis (in short AICov)''
  in the abstract. We also defined COVID-10 in the abstract.}

\REVIEW{DONE}{Acronyms need to be defined at their first occurrences
  (e.g., AI in the intro). An acronyme table is included in table 1,
  we added the following acronyms to the paper}
  {All acronyms have been defined at their first usage. In addition
    we have added a section Acronyms in which all acronyms are listed
    to make it easier for the reader to look them up in case they are
    used later on in the paper.

    All riskfactors have been capitalized.

TODO. PERCENT, PERCENTBALCK, BLACK PERCENT, also for hispanics and senior

TODO. use risk factor in table headings

TODO. PER1000 is /1000

TODO. make acronyms appear in text

TODO. Table 1. capitalization,explain why they are different or make them the same
  }

\REVIEW{TODO}{Figure 2: please present the equations as equations. The
  notations used in the equation need a full check (e.g., h is used as
  both the dimension and vector). I'd like to see a clear distinction
  between inputs/outputs and parameters. I'd suggest that the inputs
  be set up before the equations are presented.}
  { todo}

\REVIEW{TODO}{Regenerate the figures into pdf format (the current
  jpg/png figures blur when zoomed in).}
  { todo}

\REVIEW{DONE}{Table 2: consider reformating it; it would not fit the
  page width in print.}
  { Table 2 has been reformatted. It was previously fit in the column
    with resize box. While rotating the headers the font is
    increased. The table still fits in the column as it still uses
    resizebox.}

\REVIEW{DONE}{Table 3: tidy the headers.}
  { The headers have been cleaned up and linebreaks wer introduced so
    the resizebox of the table presents the table now larger} }

\REVIEW{TODO}{Figure 5-6: can be combined to one figure with two
  panels (using ggplot2 and save the space of the labels of the
  horizontal axis).  }
  { todo}

\REVIEW{TODO}{Figure 7-8: can be one figure with two panels; pop
  density 2010 needs to be cleaned.}
  { todo}

\REVIEW{TODO}{Figure 9-14: combine into one figure.}
  { todo}

\REVIEW{DONE}{Table 4: keep the number of digits consistent.
   The digits have been padded with the missing 0 to make them all
    the same length}
  { todo}

\REVIEW{TODO}{Reproducibility: submit data/code/README in a
  compressed file as supplementary material, or put them in a public
  repo and give the link under an unnumbered section Supplementary
  Material.}
  { todo}

@@

To submit your revision, please log in to EJMS and submit it as a
revised file to the original submission. Please also include a
detailed description of how you addressed all the points raised by the
reviewers.

\section*{Reviewer 2}

This paper presents AICov, an integrated deep learning framework for
COVID-19 forecasting. It takes into account population covariates, and
uses LSTM and event modeling to do the forecasting. Data integration
is done at a local level. Integration of such information brings
improvement to the prediction accuracy. Despite the authors’ effort to
address my comments in their revision, the paper is still in rough
shape. Please see my comments.

\bigskip

\REVIEW{DONE}{Page 5, please write the full name for REST
  (representational state transfer) when it first appears.}
  { The term has been defined at first usage, we also checked all
    other abbreviations. }

\REVIEW{TODO}{Page 5, line 3 under Requirement 3, “we are developed an
  abstraction API for data” does not read right.
}
  { todo}

\REVIEW{TODO}{Page 6, line 8, what level is “local level”?
}
  { todo}

\REVIEW{PROGRESS} Page 7, Table 1, is “CHD” also a percent? If
  yes please modify the description; Percent- blacks and black percent
  both correspond to “Percent of blacks in the population”. Both seem
  to have been used according to Table 3, with difference in the four
  model performance measurements. Do the authors have an explanation
  for this?
}
  { The table has been updated and it was indicated that it is in
    percent. There were indeed to different information sources that
    reported different values for percentage of blacks and other
    populations. This however has since than corrected and we have
    been able to use just one population value.

    \todo{fix the values and only have one population value}}

\REVIEW{TODO}{Page 8, line 5 of Section 5.1, the abbreviation LSTM has
  already appeared before, and the full name can be removed.
}
  { todo}

\REVIEW{TODO}{What do the two panels in Figure 4 represent? Please give
  clearer description either in the text or figure caption.
}
  { todo}

\REVIEW{TODO}{Section 5.2, paragraph 2 line 6, no need to capitalize for
  the word “Samples”. Line 5 from the bottom, no need to capitalize
  for the word “Insurance”. Please be careful with your writing.
}
  { todo}

\REVIEW{TODO}{Section 5.2, the “order for cases sorted by minimum model
  order” and “order for deaths sorted by minimum model error”
  paragraphs are redundant, as the same information is already
  displayed in the x-legend for Figures 5 and 6.
}
  { todo}

\REVIEW{TODO}{In Table 3, what is the unit for the error terms? The
  errors in Figures 7 and 8 are percentages. If errors in Table 3 are
  also percentages please clearly mark. Also, the numbers in Figures 7
  and 8 do not align with numbers in row 1 of Table 3. Please check
  your numbers with care.
}
  { todo}

\REVIEW{TODO}{Please adjust the location of your figures so they do not
  look pages apart from their interpretations.
}
  { todo}

\REVIEW{TODO}{If I am reading it correctly, Figures 9 and 11 are only
  the “zoom in” versions of partial information from Figure
  13. Similarly for the right column. Why waste space on two more
  figures when you could make Figures 13 and 14 bigger and explain the
  findings?
}
  { todo}

\REVIEW{TODO}{The last paragraph of Section 5.2 suits the
  conclusion/discussion for future work section, not in the middle of
  a paper.
}
  { todo}

\REVIEW{TODO}{Section 5.3, the authors mentioned two way of constructing
  covariates. Since the second way was more complicated while yielding
  no better performance the authors did not use it. Then it can be
  removed without damaging the storyline.
}
  { todo}

\REVIEW{TODO}{The description “We renormalized, so the two basic
  features (following day) contributed 50\% of the loss function” is
  hard to interpret. Please elaborate.
}
  { todo}

\REVIEW{TODO}{Paragraph 2 on page 14, “Selu, Relu and Tanh activations
  gave very similar results while Sigmoid activation was much
  worse. However we still use a Sigmoid function...” is very
  counter-intuitive. Isn’t the point of trying different activation
  functions to better fit the data?  16. Paragraph 3 on page 14, “The
  basic COVID-19 daily data is extensive but compared two approaches
  with and without renormalization, but in this case, we ...” The
  sentence does not read correct. Please check your sentences with
  care!
}
  { todo}

\REVIEW{TODO}{How to interpret the “significant positive correlation” in
  Section 6 from the practical perspective? Also, the term
  “evolution function” made its first appearance in Section 6 without
  any definition or explanation; I suggest either removing the whole
  Section 6, or extend it with clearer definition of evolution
  function.  18. The first two paragraphs do not read like to be in a
  “Conclusion” section. Rather, they better fit into the introduction.
}
  { todo}

### 2020-12-19 vonLaszewski-covid.tex
update todo list and abbreviations

\newcommand{\ABSTRACT}{
The COVID-19 pandemic has had profound global consequences on health,
economic, social, behavioral, and almost every major aspect of human
life. Therefore, it is of great importance to model COVID-19 and other
pandemics in terms of the broader social contexts in which they take
place.  We present the architecture of an artificial intelligence
enhanced COVID-19 analysis (in short AICov) , which provides an
integrative deep learning framework for COVID-19 forecasting with
population covariates, some of which may serve as putative risk
factors.  We have integrated multiple different strategies into AICov,
including the ability to use deep learning strategies based on LSTM
and event modeling.  To demonstrate our approach, we have introduced a
framework that integrates population covariates from multiple
sources. Thus, AICov not only includes data on COVID-19 cases and
deaths but, more importantly, the population's socioeconomic, health,
and behavioral risk factors at a local level. The compiled data are
fed into AICov, and thus we obtain improved prediction by the
integration of the data to our model as compared to one that only uses
case and death data. As we use deep learning our models adapt over
time while learning the model from past data.
}

@@

In this study, we used a Long Short-Term Memory (LSTM) algorithm for
our predictions \citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a
recurrent neural network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

@@

\begin{requirement}{\bf Ease of Use.}
  A major design criterion for AICov is that the framework must be
  easy to use, and extensions can be made to allow usability of both
  the API as well as the user interface. \Solution To address these
  requirements, we are devising a specialized but easy to use time
  series API for COVID-19 that integrates common tasks such as
  automated filtering and normalization of the data.
\end{requirement}

@@

\begin{requirement}{\bf Compute Resource Mapping.}
  For other uses of AICov, we need to integrate a flexible resource
  utilization framework. \Solution To address this requirement, we
  leverage from our earlier work and utilize through Cloudmesh and
  NIST NBDIF definitions cloud services into the architecture
  \citep{las-19-nist}. This includes containers but also infrastructure
  services in the cloud. Through these service integrations, we also
  provide access to GPUs.
\end{requirement}

@@

While the U.S. has seen an overall case fatality rate of 5.14\%, some
U.S. locations have experienced a disproportionate number of deaths
compared to others. This disparity of deaths among different
communities could be attributed to a diversity of local risk factors
and socioeconomic determinants of health that now constitute a very
active area of investigation. Researchers are studying the effects of
different community-specific socioeconomic variables, underlying
health conditions and comorbidities, infrastructure, and systemic
responses.  Given the complexity of such underlying conditions that
can determine individual disease outcomes, where and under what
conditions people get infected by the virus, and then recover or die
take place not in isolation but rather in an interplay of diverse
factors that are characteristic to a local community, which need to be
synthesized. In other words, a basic requirement for effective
modeling of COVID-19 dynamics is that the information regarding a
particular community’s vulnerabilities and strengths pertaining to the
disease must be as comprehensive and integrative as possible.  The
regular reporting of accurate Coronavirus outbreaks data from local
health departments has been understandably difficult, especially in
the areas that are hardest hit. We use data on community-specific
health and behavioral risk factors aggregated at the county level from
the large-scale longitudinal CDC BRFSS surveys \citep{www-cdc-brfss},
and demographic and socioeconomic data from the U.S. Census Bureau
that are known at the local levels along with the cases and deaths
data compiled by NCHS \citep{www-cdc-nchs} and integrate them with
other data sources.

@@

Among the different sources of data on daily cases and deaths data
collected and aggregated at different spatial levels, especially for
the U.S., many are either freely available repositories such as GitHub
or accessible via specifically developed online resources such as the
New York Times~\citep{github-nytimes} and the Johns Hopkins Coronavirus
Resource Center~\citep{www-jh-covid19}.  For this study, we generated
two data sets based on the latter resource upon our checking for the
locations and times at which cumulative counts of COVID-19 cases and
deaths on any given day were no less than those on the previous
day. This led us to compile daily coronavirus cases and deaths data
for 345 U.S. counties from 32 states and the District of Columbia and
matched by five-digit FIPS code or county name to dynamic and static
variables from additional data sources.

@@

The final choice was the quantities to predict and train on. We choose
this to be both the cases and deaths values at the time value
following the sequence combined with the same quantities for the
following 14 days. This gives us 30 different output features for each
sequence. We renormalized, so the two basic features (following day)
contributed 50\% of the loss function. We supported this in TensorFlow
with a custom mean square error ($MSE$) loss function that dropped
training values that were not available for time sequences near the
end of the examined period. While one can not train on future values,
we can predict them, as seen in our sample outputs. We extensively
studied both 2 and 30 output features, and they give similar
descriptions of the observed data with the 30 feature choice enabling
14-day predictions.

We performed an extensive hyper-parameter search and we show the
values in Table \ref{tab:model} that give robust results without
overfitting. Furthermore, we looked at four different activation
functions for this analysis in the LSTM and fully connected
layers. $SELU$, $RELU$ and $TANH$ activations gave very similar
results while $SIGMOID$ activation was much worse. However we still
use a $SIGMOID$ function in the last stage just as we did in Section
\ref{sec:lstm-covariate}. We also looked at the number of hidden units
(separately for LSTM and fully connected layers), the number of LSTM
layers and the dropout values.  Only the latter produced a significant
change and, for example, removing dropouts decreased the loss by about
16\% but we retained them to provide a robust environment. We chose
randomly 20\% of the data for validation and testing and this had
similar loss values to the training data (the other 80\%) with our
chosen dropout.

The final technical issue concerns the data normalization where each
feature is transformed by a feature dependent -- but city independent
-- way to lie between 0 and 1. Further all extensive (proportional to
size) covariates were first divided by the city population except for
the population itself where we took the logarithm; intensive
covariates were just rescaled to lie between 0 and 1. The basic
COVID-19 daily data is extensive but compared two approaches with and
without renormalization, but in this case, we took the square root of
the data value.  This is natural if there are counting (square root(N)
for N observations) errors as the deep learning loss function is
$(predicted-observed)^{2}$ without the error traditional in
$\chi^{2}$, namely $((predicted-observed)/error)^{2}$. This approach
with standard deep learning $MSE$ appears sound instead of dividing by
the population. That will emphasize the error contribution of smaller
cities but is still interesting to look at due to the presence of many
systematic errors.

@@

We present AICov as an integrative deep learning framework for
COVID-19 forecasting with the help of population covariates. Thus, its
objectives are different from those of traditional compartmental
models. One of the important features of the architecture for AICov is
that it is by design targeting Cloud and HPC resources to conduct
parameter sweeps to leverage sophisticated deep learning toolkits. The
architecture allows the integration of various data sources that can
update the data from its sources on demand and update its model
predictions based on newly introduced data. Parameters can easily be
adjusted via Jupyter notebooks that, in turn, call the computational
backends on HPC and cloud resources.

@@

This work is partially supported by the National Science Foundation
(NSF) through awards CIF21 DIBBS 1443054, nanoBIO 1720625,
Cybertraining 1829704, CINES 1835598 and Global Pervasive
Computational Epidemiology 1918626.  We thank J. Kadupitiya for
several useful discussions on deep learning frameworks.

@@

This section contains the list of acronyms in Table{tab:acronymes} as used in the paper for easy
reference. The definitions of the risk factors are however given
within the paper in Table~\ref{tab:risk-factors} on page
\pageref{tab:risk-factors} and not repeated here.

\begin{table}[!h]
  \begin{tabular}{ll}
    \toprule
    Abbreviation & Description \\
    \midrule
    API & Application Programming Interfaces \\
    BRFSS & Behavioral Risk Factor Surveillance System \\
    CDC & Center for Desease Control \\
    CI & Cyber Infrastructure \\
    CIF21 & Cyberinfrastructure Framework for 21st Century  \\
    CINES & CyberInfrastructure for Network Engineering and Science \\
    COVID-19 & COrona VIrus Disease 2019  \\
    CSSE & Center for Systems Science and Engineering \\
    cum & cumulative \\
    DIBBS & Data Infrastructure Building Blocks \\
    FIPS & Federal Information Processing Standards \\
    GPU & Graphics Processing Unit  \\
    HPC & High Performance Computing \\
    LSTM & Long Short-Term Memory \\
    MSE & Mean Squared Error \\
    NBDIF & NIST Big Data Working Group \\
    NCHS & CDC National Center for Health Statistics \\
    NIST & National Institute of Standards and Technology \\
    NONE & No risk factors used \\
    NORM\_POP & Normalized Population \\
    NSF & National Science Foundation \\
    RELU & REctified Linear Unit \\
    REST & REpresentational State Transfer  \\
    RNN & Recurrent Neural Network \\
    SEIR & Susceptible, Expose, Infectious, Recovered) \\
    SELU & Scaled Exponential Linear Unit \\
    SIR &  Susceptible, Infectious, Recovered \\
    U.S. & United States of America \\
    WHO & World Health Organization \\
    \midrule
\end{tabular}
\end{table}

>>>>

\newcommand{\ABSTRACT}{ The COVID-19 (COrona VIrus Disease 2019)
  pandemic has had profound global consequences on health, economic,
  social, behavioral, and almost every major aspect of human
  life. Therefore, it is of great importance to model COVID-19 and
  other pandemics in terms of the broader social contexts in which
  they take place.  We present the architecture of an artificial
  intelligence enhanced COVID-19 analysis (in short AICov) , which
  provides an integrative deep learning framework for COVID-19
  forecasting with population covariates, some of which may serve as
  putative risk factors.  We have integrated multiple different
  strategies into AICov, including the ability to use deep learning
  strategies based on Long Short-Term Memory (LSTM) and event
  modeling.  To demonstrate our approach, we have introduced a
  framework that integrates population covariates from multiple
  sources. Thus, AICov not only includes data on COVID-19 cases and
  deaths but, more importantly, the population's socioeconomic,
  health, and behavioral risk factors at a local level. The compiled
  data are fed into AICov, and thus we obtain improved prediction by
  the integration of the data to our model as compared to one that
  only uses case and death data. As we use deep learning our models
  adapt over time while learning the model from past data.}

@@

In this study, we used a Long Short-Term Memory (LSTM) algorithm for
our predictions \citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a
Recurrent Neural Network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

@@

\begin{requirement}{\bf Ease of Use.}
  A major design criterion for AICov is that the framework must be
  easy to use, and extensions can be made to allow usability of both
  the Application Programming Interfaces (API) as well as the user
  interface. \Solution To address these requirements, we are devising
  a specialized but easy to use time series API for COVID-19 that
  integrates common tasks such as automated filtering and
  normalization of the data.
\end{requirement}

@@

\begin{requirement}{\bf Compute Resource Mapping.}
  For other uses of AICov, we need to integrate a flexible resource
  utilization framework. \Solution To address this requirement, we
  leverage from our earlier work and utilize through Cloudmesh and
  National Institute of Standards and Technology (NIST) Big Data
  Working Group (NBDIF) definitions cloud services into the
  architecture \citep{las-19-nist}. This includes containers but also
  infrastructure services in the cloud. Through these service
  integrations, we also provide access to Graphics Processing Units
  (GPUs).
\end{requirement}

@@

While the U.S. has seen an overall case fatality rate of 5.14\%, some
U.S. locations have experienced a disproportionate number of deaths
compared to others. This disparity of deaths among different
communities could be attributed to a diversity of local risk factors
and socioeconomic determinants of health that now constitute a very
active area of investigation. Researchers are studying the effects of
different community-specific socioeconomic variables, underlying
health conditions and comorbidities, infrastructure, and systemic
responses.  Given the complexity of such underlying conditions that
can determine individual disease outcomes, where and under what
conditions people get infected by the virus, and then recover or die
take place not in isolation but rather in an interplay of diverse
factors that are characteristic to a local community, which need to be
synthesized. In other words, a basic requirement for effective
modeling of COVID-19 dynamics is that the information regarding a
particular community’s vulnerabilities and strengths pertaining to the
disease must be as comprehensive and integrative as possible.  The
regular reporting of accurate Coronavirus outbreaks data from local
health departments has been understandably difficult, especially in
the areas that are hardest hit. We use data on community-specific
health and behavioral risk factors aggregated at the county level from
the large-scale longitudinal CDC Behavioral Risk Factor Surveillance
System (BRFSS) surveys \citep{www-cdc-brfss}, and demographic and
socioeconomic data from the U.S. Census Bureau that are known at the
local levels along with the cases and deaths data compiled by the CDC
National Center for Health Statistics (NCHS) \citep{www-cdc-nchs} and
integrate them with other data sources.

@@

Among the different sources of data on daily cases and deaths data
collected and aggregated at different spatial levels, especially for
the U.S., many are either freely available repositories such as GitHub
or accessible via specifically developed online resources such as the
New York Times~\citep{github-nytimes} and the Johns Hopkins
Coronavirus Resource Center~\citep{www-jh-covid19}.  For this study,
we generated two data sets based on the latter resource upon our
checking for the locations and times at which cumulative counts of
COVID-19 cases and deaths on any given day were no less than those on
the previous day. This led us to compile daily coronavirus cases and
deaths data for 345 U.S. counties from 32 states and the District of
Columbia and matched by five-digit Federal Information Processing
Standards (FIPS) code or county name to dynamic and static variables
from additional data sources.

@@

The final choice was the quantities to predict and train on. We choose
this to be both the cases and deaths values at the time value
following the sequence combined with the same quantities for the
following 14 days. This gives us 30 different output features for each
sequence. We renormalized, so the two basic features (following day)
contributed 50\% of the loss function. We supported this in TensorFlow
with a custom mean square error (MSE) loss function that dropped
training values that were not available for time sequences near the
end of the examined period. While one can not train on future values,
we can predict them, as seen in our sample outputs. We extensively
studied both 2 and 30 output features, and they give similar
descriptions of the observed data with the 30 feature choice enabling
14-day predictions.

We performed an extensive hyper-parameter search and we show the
values in Table \ref{tab:model} that give robust results without
overfitting. Furthermore, we looked at four different activation
functions for this analysis in the LSTM and fully connected
layers. Scaled Exponential Linear Unit (SELU), REctified Linear Unit
(RELU) and TANH activations gave very similar results while SIGMOID
activation was much worse. However we still use a SIGMOID function in
the last stage just as we did in Section \ref{sec:lstm-covariate}. We
also looked at the number of hidden units (separately for LSTM and
fully connected layers), the number of LSTM layers and the dropout
values.  Only the latter produced a significant change and, for
example, removing dropouts decreased the loss by about 16\% but we
retained them to provide a robust environment. We chose randomly 20\%
of the data for validation and testing and this had similar loss
values to the training data (the other 80\%) with our chosen dropout.

The final technical issue concerns the data normalization where each
feature is transformed by a feature dependent -- but city independent
-- way to lie between 0 and 1. Further all extensive (proportional to
size) covariates were first divided by the city population except for
the population itself where we took the logarithm; intensive
covariates were just rescaled to lie between 0 and 1. The basic
COVID-19 daily data is extensive but compared two approaches with and
without renormalization, but in this case, we took the square root of
the data value.  This is natural if there are counting (square root(N)
for N observations) errors as the deep learning loss function is
$(predicted-observed)^{2}$ without the error traditional in
$\chi^{2}$, namely $((predicted-observed)/error)^{2}$. This approach
with standard deep learning MSE appears sound instead of dividing by
the population. That will emphasize the error contribution of smaller
cities but is still interesting to look at due to the presence of many
systematic errors.

@@

We present AICov as an integrative deep learning framework for
COVID-19 forecasting with the help of population covariates. Thus, its
objectives are different from those of traditional compartmental
models. One of the important features of the architecture for AICov is
that it is by design targeting Cloud and High Performance Computing
(HPC) resources to conduct parameter sweeps to leverage sophisticated
deep learning toolkits. The architecture allows the integration of
various data sources that can update the data from its sources on
demand and update its model predictions based on newly introduced
data. Parameters can easily be adjusted via Jupyter notebooks that, in
turn, call the computational backends on HPC and cloud resources.

@@

This work is partially supported by the National Science Foundation
(NSF) through awards Cyberinfrastructure Framework for 21st Century
Data Infrastructure Building Blocks (1443054), Network for
Computational Nanotechnology Engineered nanoBIO Node (1720625),
Cybertraining (1829704), CyberInfrastructure for Network Engineering
and Science (1835598) and Global Pervasive Computational Epidemiology
(1918626).  We thank J. Kadupitiya for several useful discussions on
deep learning frameworks.

@@

This section contains the list of acronyms in Table \ref{tab:acronyms}
as used in the paper for easy reference. The definitions of the risk
factors are however given within the paper in
Table~\ref{tab:risk-factors} on page \pageref{tab:risk-factors} and
not repeated here.

\todo{NORM\_POP Normalized population}

\begin{table}[!h]
  \caption{Non-risk factor acronyms used in this paper.}
  \label{tab:acronyms}
  \begin{tabular}{ll}
    \toprule
    Abbreviation & Description \\
    \midrule
    API & Application Programming Interfaces \\
    BRFSS & Behavioral Risk Factor Surveillance System \\
    CDC & U.S.~Centers for Disease Control and Prevention \\
    CI & Continious INtegration \\
    COVID-19 & COrona VIrus Disease 2019  \\
    CSSE & Center for Systems Science and Engineering \\
    cum & cumulative \\
    FIPS & Federal Information Processing Standards \\
    GPUs & Graphics Processing Units  \\
    HPC & High Performance Computing \\
    LSTM & Long Short-Term Memory \\
    MSE & Mean Squared Error \\
    NBDIF & NIST Big Data Working Group \\
    NCHS & CDC National Center for Health Statistics \\
    NIST & National Institute of Standards and Technology \\
    NONE & No risk factors used \\
    NORM\_POP & Normalized Population \\
    NSF & National Science Foundation \\
    RELU & REctified Linear Unit \\
    REST & REpresentational State Transfer  \\
    RMSE & Root Mean Squared Error \\
    RNN & Recurrent Neural Network \\
    SEIR & Susceptible, Expose, Infectious, Recovered) \\
    SELU & Scaled Exponential Linear Unit \\
    SIR &  Susceptible, Infectious, Recovered \\
    U.S. & United States of America \\
    WHO & World Health Organization \\
    \midrule
  \end{tabular}
\end{table}
